Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2023-02-18 analysis/paper/paper.qmd
Done with dealing with BM's comments except for adding CV and ANOVA

Our results indicate that javelin tips and stabbing spear tips are the most probable ballistic uses for stemmed points. There are a few stemmed points in our sample that may have been used as dart tips and un-poisoned arrow tips. None have TCSA values in the range of poison arrow tips (@fig-tcsa-all-sp). In their study, @lee2019 calculated TCSA values for ten stemmed points from Jingeuneul site, which were not included in the research due to data availability issues, and concluded that these values are within the range of North American dart tips and arrowheads. In general, we found a wide range of TCSA in our sample, suggesting that stemmed points might play diverse roles in foraging toolkits.

TCSA can be impacted by other factors such as raw materials and portability. @eren2022plains explain that smaller TCSA could be the result of pursuing production economy and transport efficiency. We explored the relationship between TCSA and raw material types and weight in order to examine these factors. We observed no clear pattern for raw materials (@fig-tcsa-raw-materials) but weight (@fig-artifact-size-cluster), though weight is correlated with raw materials (@fig-tcsa-size). We speculate that this might be due to the higher availability of raw materials, such as shale and hornfels on the landscape. Therefore, we could assume that raw materials also associated with TCSA value. We found that different clusters of weight are matched with different types of weapon tips (@fig-artifact-size-cluster).

>>>>

Our results indicate that javelin tips and stabbing spear tips are the most probable ballistic uses for stemmed points. There are a few stemmed points in our sample that may have been used as dart tips and un-poisoned arrow tips. None have TCSA values in the range of poison arrow tips (@fig-tcsa-all-sp). In their study, @lee2019 calculated TCSA values for ten stemmed points from Jingeuneul site, which were not included in the research due to data availability issues, and concluded that these values are within the range of North American dart tips and arrowheads. We found a wide range of TCSA in our sample covering the range of dart tips and un-poisoned arrow tips, which are assumed to be equivalent to arrowheads. In general, the wide range of TCSA suggest that stemmed points might play diverse roles in foraging toolkits.

TCSA can be impacted by other factors such as raw materials and portability. @eren2022plains explain that smaller TCSA could be the result of pursuing production economy and transport efficiency. We explored the relationship between TCSA, and weight and raw material types in order to examine these factors. We found that different clusters of weight are matched with different types of weapon tips (@fig-artifact-size-cluster). There was no clear pattern for raw materials (@fig-tcsa-raw-materials). However, combined with weight, we observed a positive relationship between raw material type and TCSA values (@fig-tcsa-size). We speculate that this might be due to the higher availability of raw materials, such as shale and hornfels on the landscape. Therefore, we could assume that raw materials also associated with TCSA value.
